Designing Modular Software: A Case Study in Introductory Statistics

### 2016-10-10 jcgs-paper-2016/paper.Rmd
Small fixes

The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community. However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student [@5359977]. This additional time commitment may explain why introductory courses often utilize point-and-click applications, even if the instructor himself/herself uses R in their own work. Still, some compromises must be made when using many graphical applications, including dealing with software licenses and unsupported desktop platforms.

@@

In the development of `intRo`, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find what they need more easily. We also made the functionality available the minimal necessary for an introductory statistics course. Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in `intRo`. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/user_experience.pdf}
\caption{A typical student experience of generating a result in \texttt{intRo}. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
\label{fig:user_experience}
\end{figure}

@@

On the front end, user interaction with `intRo` is split into bitesize chunks that we call modules. Modules are self-contained pieces of functionality which implement common statistical procedures that are often used in introductory statistics classes. These modules form the core of `intRo` and are discussed at length in the next section.

@@

`intRo` modules are not to be confused with Shiny modules [@shinymodules]. Shiny modules are a recent feature added to Shiny which allows the bundling of inputs and outputs into a single set of functionality. They are more general and suitable for any application. `intRo` modules are specifically for statistics functionality and work within the `intRo` application only.

@@

The modules provided with `intRo` are contained in the modules folder. The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `inference`). Within each of these categories is a folder named according to the name of the module. This folder houses the previously defined scripts. As an example, we will walk through the contents of the `nonparametric` module as previously mentioned in this section.

Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence can be found in the `intRo` repository at **modules/inference/nonparametric**. Let's first examine *helper.R*:

>>>>

The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community. However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student [@5359977]. This additional time commitment may explain why introductory courses often utilize point-and-click applications, even if the instructor himself/herself uses R in their own work. Still, some compromises must be made when using many graphical applications, including dealing with software licenses and unsupported desktop platforms. From the instructor's perspective in particular, managing a large group of software licenses for students with various computing environments and versions could wind up being extremely cumbersome.

@@

In the development of `intRo`, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find what they need more easily. We also made the functionality available the minimal necessary for an introductory statistics course. Figure \ref{fig:user_experience} presents a schematic of the simple steps a student takes to generate a result in `intRo`. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/user_experience.pdf}
\caption{A schematic of the typical student experience of generating a result in \texttt{intRo}. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
\label{fig:user_experience}
\end{figure}

@@

On the front end, user interaction with `intRo` is split into bitesize chunks that we call modules. In `intRo`'s context, modules are self-contained pieces of functionality which implement common statistical procedures. These modules form the core functionality of `intRo` and are discussed at length in the next section.

@@

`intRo` modules are not to be confused with Shiny modules [@shinymodules]. Shiny modules are a recent feature added to Shiny which allows the bundling of inputs and outputs into a single set of functionality. They are more general and suitable for any application. `intRo` modules are for statistics functionality and work within the `intRo` application only.

@@

The modules provided with `intRo` are contained in the modules folder. The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `inference`). Within each of these categories is a folder named according to the name of the module. This folder houses the previously defined scripts. As an example, we will walk through the process of creating a new module called `nonparametric`, as previously mentioned in this section, which will perform a wilcoxon rank sum test.

Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence would be placed in the `intRo` repository at **modules/inference/nonparametric**. Let's first examine *helper.R*:
